---
title: "similaRpeak: Using metagene output to compare profiles"
author: "Charles Joly Beauparlant, Fabien C. Lamaze, Astrid L. Deschênes, 
Rawane Samb, Elsa Bernatchez, Pascal Belleau, Steve Bilodeau and Arnaud Droit"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{similaRpeak: Using metagene output to compare profiles}
  %\VignetteEngine{knitr::rmarkdown}
  \usepackage[utf8]{inputenc}
---

<center> <h1> Enrichment of the H3K27ac histone mark in active and bivalent transcription start sites (TSS) </h1> </center>

## Introduction

The H3K27ac histone mark is normally enriched at active transcriptional start 
sites (TSS) compared to poised TSS (Roadmap Epigenomics Consortium et al.,2015).
_metagene_  has been used to robustly test for the difference in enrichment of 
H3K27ac between these two groups of promoter regions in A549 cells. 


### 1. Load _metagene_ results data

The results have been saved in a `list` that can be loaded and used by 
`similaRpeak`. The `list` contains 2 elements: a `data.frame` and a Friedman test 
results. Only the `data.frame` is needed.

```{r loadData}
load("../data/df.RData")
DF<-df[["DF"]]
```

The `data.frame` content can be used to recreate the _metagene_ graph.

```{r graph, fig.height=5, fig.width=7, fig.align='center', echo=FALSE}
library(ggplot2)
p <- ggplot(DF, aes(x=position, y=value, ymin=qinf, ymax=qsup))
p + geom_ribbon(aes(fill=group), alpha=0.3) + 
    geom_line(aes(color=group), size=1, litype=1, bg="transparent") +
    ylab("Reads per million (RPM)") 

```

### 2. Load _similaRpeak_ package

The _similaRpeak_ package must be loaded.

```{r loadPackage, warning=FALSE, collapse=TRUE}
library(similaRpeak)
```


### 3. Calculate metrics using _similaRpeak_

The metrics are calculated using the _similarity_ function which takes 
as arguments the two profiles vectors and the threshold values.

```{r metrics, collapse=TRUE}
# H3K27ac_TssA vector
H3K27ac<-DF$value[DF$group == "H3K27ac_TssA"]
# H3K27me3_TssA vector
H3K27me3<-DF$value[DF$group == "H3K27me3_TssA"]
# Calculate metrics
metrics <- similarity(H3K27ac, 
                            H3K27me3, 
                            ratioAreaThreshold=0.1, 
                            ratioMaxMaxThreshold=0.01, 
                            ratioIntersectThreshold=0.1, 
                            ratioNormalizedIntersectThreshold=0.1,
                            diffPosMaxThresholdMinValue=0.01, 
                            diffPosMaxThresholdMaxDiff=100, 
                            diffPosMaxTolerance=0.01)
metrics
```

The `RATIO_AREA` and `RATIO_MAX_MAX` reflect the difference of amplitude 
between both profiles.  The small `RATIO_INTERSECT` is caused by the small
values of the H3K27me3 profile. 

The `RATIO_NORMALIZED_INTERSECT` is quite high. However, the graph of the
normalized histone profiles shows that this value is caused by the relatively
flat profiles.

```{r graphNorm, echo=FALSE, fig.align='center', fig.height=5, fig.width=7}
plot(H3K27ac*length(H3K27ac)/sum(H3K27ac, na.rm=TRUE), xaxt = "n", 
        type="l", col="blue", xlab="", ylab="", ylim=c(0.5, 2.5))
par(new=TRUE)
plot(H3K27me3*length(H3K27me3)/sum(H3K27me3, na.rm=TRUE), 
        type="l", col="darkgreen", xlab="position", 
        ylab="Normalized Coverage (Coverage/Mean Coverage)", 
        ylim=c(0.5, 2.5), xaxt = "n")
legend("topright", c("H3K27ac_TssA","H3K4me1_TssA"), cex=1.2, 
        col=c("blue","darkgreen"), lty=1)
```

In conclusion, the profile of H3K27ac histone mark in active transcriptional 
start sites (TSS) compared to poised TSS are not sharing any of the tested 
patterns.

## References 

Roadmap Epigenomics Consortium et al. (2015). Integrative analysis of 111 
reference human epigenomes. Nature, 518: 317–330.

